Title here

### 2016-12-08 portfolio/text/ms.Rmd
Edit

Managing natural resources and sustaining the livelihoods of those that depend
on their extraction is difficult. These professions tend to have high
year-to-year variability in income relative to individuals who don't rely on
natural resources as a source of income  [@mishra2002; @sethi2012b]. In
addition to economic risk, such as changes in demand or prices, individuals
dependent on natural resources are also subject to biological and environmental
variability. For example, drought and flooding are a major source of risk to
agricultural food security and farmers' incomes [@gregory2005], and
catastrophic disease outbreaks and wildfires increase risk for the logging
industry [@kirilenko2007].

Individuals who rely on natural resources for income develop strategies to
reduce income variability. For example, farmers may rotate their crops to buffer
short-term climate variation and long-term environmental trends [@zentner1988].
Similarly, loggers maintain a diverse mix of tree species, age structure, and
harvest in geographical areas to buffer against disease outbreak, wildfires, and
price variation REF. However, external forces may restrict the options available
to individuals to reduce variability. For example, otherwise well-intentioned
regulations may limit how individuals diversify, or incentivize against
diversification. For example, in terrestrial agriculture, crop subsidies in the
US may incentivize some farms to specialize (e.g. dairy, grains, soybeans;
@mishra2002). In other instances, external market forces (prices, demand) may
discourage the adoption of diverse strategies and help promote specialization
[@bradshaw2004].

Commercial fishers present an interesting example of diversification, given the
extreme risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
available to terrestrial farmers (e.g. crop insurance programs) have not been
obtainable by commercial fishers [@kasperski2013], possibly because both fish
and fishers are spatially mobile. However, fishers still have the ability to
lower risk. They can, for example, target as many species as permitted by
a license, they may own multiple licenses, or diversify spatially and fish in
multiple regions (increasing their total catch, or numbers of species caught).
The effectiveness of these diversification strategies for improving revenue and
reducing revenue variability remains uncertain. Furthermore, theory and
empirical studies both suggest there can be substantial benefits to
specialization for both farmers and fishers [@abdulai2000; @arnason2005]. The
majority of papers that examine effects of diversification on revenue do so at
aggregate levels that combine individuals (e.g. households, counties,
communities; @sethi2014; @kasperski2013). Because of this, key uncertainties
remain about the fundamental relationship between diversity, fishing revenue,
and fishing revenue variability at the level of the individual.

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability. Fisheries in Alaska are among the highest-valued fisheries in the
world with 2.7 billion kg landed biomass valued at nearly 2 billion USD annually
[@nmfs2016]. The industry supports a range of fishing fleets ranging from small,
open, single-operator vessels to multi-million-dollar state-of-the-art ships
with large crews. Like much of the rest of the world [@hilborn2003a], the days
of open, unregulated fisheries in most developed countries have passed ---
fishers must own a license to participate in virtually all fisheries. And while
such regulations have proven helpful in reducing or eliminating overfishing,
they limit the available strategies to reduce income variability. For example,
many permits limit fishing to a single species, diversifying by purchasing
additional fishing permits can be expensive (e.g. ~\$192K USD for a king and
tanner crab permit in Southeast Alaska in 2016;
<https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>), and some fisheries limit the
entry of new participants through individual fishing quotas (IFQs) [@grimm2012].

Using a unique dataset collected over 30 years (1985--2014) on individuals,
rather than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals rather than
vessels we can track people that switch fishery sectors over time, or alter
strategies and use multiple vessels. Similarly, we are able to disentagnle
fishers that share the same vessel. We develop and apply a hierarchical Bayesian
model to understand the effect of diversity on both individual revenue and
revenue variability that builds on results from vessel [@kasperski2013], fleet
[@sethi2012b], and community-level [@sethi2014] variability. Our model controls
for fishing effort and latent time effects (e.g. changing environmental
conditions), simultaneously estimates the effect of species diversity on
individual revenue and revenue variability, and estimates effects both within
and across fishing strategies (sets of similar licenses or permits). To our
knowledge, this is the first application of this type of model in any natural
resource field.

# Methods

We developed a hierarchical variance-function regression model [@davidian1987;
@hall1989], which separately but simultaneously models changes in the mean and variance of
individual revenue as a function of covariates. Whereas traditional
hierarchical linear models assume homoscedasticity --- that the residual
variance does not vary systematically --- the variance-function regression model
allows the variance to be heteroscedastic and vary as a function of predictors.
Our specific model allows certain fishing strategies and levels of species
diversification to generate more or less revenue and be more or less variable
than others.

## Data

We used combinations of state and federal fishing permits held by an individual
in in a specific year to define fishing "strategies", representing the species
targeted (Supporting Information). For example, we combined permits that were
otherwise the same except for vessel size (e.g. a longliner fishing sablefish
on a vessel under 60 ft was considered to have the same strategy as a longliner
fishing sablefish on a vessel over 60 ft), and we combined a number of permits
that were targeting a single species and differed only in region fished.
Therefore, for example, someone fishing herring roe in Kodiak and someone
fishing herring roe in Alaska Peninsula were both considered to have a 'herring
roe' permit group. For salmon permits, we formed salmon strategies based on
similarities in gear type and species compositon (Supporting Information).

\begin{equation}
\log(R_{i,t}) \sim \mathrm{N} \left(\widehat{\log(R_{i,t})}, \widehat{\sigma_{i,t}^2} \right).
\end{equation}

\noindent Each year $t$ of revenue for fisher $i$ is assigned a strategy indexed
by $j$ (Supporting Information, Table \ref{tab:strategies}) and expected
$\log(R_{i,t})$ is modelled as:

<!-- Ole: I might swap out log(R)-hat for something like mu(or your favorit greek letter) just to avoid -->
<!-- potential confusion like using the term "expected log(R)".   -->
<!-- I also just don't like hats... but that's my problem. -->

\begin{equation}
\label{eq:mean}
\widehat{\log(R_{i,t})} = \beta_{0,j,t} -
  \beta_{1,j} S_{i,t} I_{i,t} +
  \beta_{2,j} S_{i,t} I_{i,t} +
  \beta_3 D_{i,t} -
  \beta_4 S_{i,t} D_{i,t} I_{i,t} +
  \beta_5 S_{i,t} D_{i,t} I_{i,t} +
  \log(R_{i,t-1}),
\end{equation}

\noindent where $S_{i,t}$ is the percent change in species diversity, $\log
\left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$, $D_{i,t}$ is the percent
change in days fished, $\log
\left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and $I_{i,t}$ is an
indicator variable that takes a value $0$ if $\mathrm{div}_{i,t} \ge
\mathrm{div}_{i,t-1}$ (species diversity increasing; generalizing) and takes
a value $1$ if $\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$ (species diversity
decreasing; specializing). Estimating a separate coefficient for the effect of
increasing diversity and decreasing diversity allows the costs and benefits of
these changes to be asymmetric.

We use $\log(R_{i,t-1}$) as an offset, and thus our model describes the percent
change in revenue, i.e., $\log(R_{i,t}) - \log(R_{i,t-1})
= \log(R_{i,t}/R_{i,t-1})$. We model change in revenue because preliminary
analysis revealed non-stationary trends in revenue (resulting from experience or
other unmodeled factors). This approach eliminates having to deal with the
non-stationary time series and eliminates having to estimate thousands of
mean-revenue intercepts associated with individual fishers. We removed
person-year combinations without a subsequent year of revenue.

The $\beta$ terms represent estimated coefficients, some of which are allowed to
vary as "random effect" terms. The intercepts $\beta_{0,j,t}$ are allowed to
vary by strategy-year combinations, are centered on zero, and are constrained by
a normal distribution:

\begin{align}
\beta_{1,j} &\sim \mathrm{N} \left(\mu_{\beta_1}, \sigma^2_{\beta_1}\right),\\
\beta_{2,j} &\sim \mathrm{N} \left(\mu_{\beta_2}, \sigma^2_{\beta_2}\right).
\end{align}

\noindent The $\beta_{1,j}$'s represent how much a 1\% decrease in species
diversity within a strategy translates to a Y\% change in revenue. The
$\beta_{2,j}$'s represent how much a 1\% increase in species diversity within
a strategy translates to a Y\% change in revenue. These are after controlling
for percent change in days fished, $\beta_3$, and the interaction between days
fished and percent change in species diversity, $\beta_4$ and $\beta_5$.

\begin{equation}
\label{eq:cv}
\widehat{\sigma_{i,t}} =
\exp \left( \gamma_{0,j} -
\gamma_{1,j} S_{i,t} I_{i,t} +
\gamma_{2,j} S_{i,t} I_{i,t} +
\gamma_3 D_{i,t} -
\gamma_4 S_{i,t} D_{i,t} I_{i,t} +
\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
\end{equation}

\begin{align}
\gamma_{0,j}  &\sim \mathrm{N} \left(\eta_0 + \eta_1 M_j + \eta_2 E_j, \sigma^2_{\gamma_0}\right),\\
\gamma_{1,j}  &\sim \mathrm{N} \left(\mu_{\gamma_1}, \sigma^2_{\gamma_1}\right),\\
\gamma_{2,j}  &\sim \mathrm{N} \left(\mu_{\gamma_2}, \sigma^2_{\gamma_2}\right).
\end{align}

\noindent The mean of the $\gamma_{0,j}$ coefficients is modelled with two
strategy-level predictors. The symbols $M_j$ and $E_j$ represent the raw mean
species diversity and mean combined fishing season length (effort) for strategy
$j$. The coefficients $\eta_1$ and $\eta_2$ represent predictors in this
strategy-level regression estimating effects across strategies.

We fit our models with Stan 2.11.1 [@hoffman2014; @standevelopmentteam2015;
@standevelopmentteam2015a; @carpenter2015] and R 3.3.1 [@rcoreteam2016]. Stan
implements the No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm
[@hoffman2014] to perform Bayesian statistical inference. We assigned weakly
informative priors on all parameters: $N(0, 2^2)$ priors on all $\beta$,
$\gamma$, and $\eta$ parameters, and half-$t(3, 0, 2)$ priors (i.e. degrees of
freedom of 3 and scale of 2) on all $\sigma$ parameters to allow for heavier
tails. We ran four chains and 2,000 iterations and discarded the first 1,000
iterations of each chain as warm up (i.e. 4,000 final samples). We checked for
chain convergence visually with trace plots. We also ensured that
$\hat{R}<1.05$ (the potential scale reduction factor) and that the effective
sample size, as calculated by the rstan R package [@standevelopmentteam2015],
was greater than 200 for all parameters [@gelman2014]. We verified that our
model returned sensible estimates by comparing our estimates to an ad-hoc
two-stage model fit with the lme4 R package [@bates2015], where we first fit a
mixed effect regression model to the mean, and secondly modeled the absolute
residuals from that first model in a second mixed effects model representing
the variance component. We also validated the model with simulated data, and
were able to recover the true values (Supporting Information). Code to
reproduce our results (with a simulated data set) is available at
<https://github.com/NCEAS/pfx-commercial>.

*TODO: for the following into the results section or and the method section with
this:*

@@

We evaluated trends in revenue and diversification over time for individuals
fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
majority (70--90%) of fishers hold single permits, with a slight increase in
single-permit holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is
most sensitive to halibut or salmon permit holders --- compared to other
fisheries, halibut permit holders are more likely to hold halibut plus other
permits, whereas salmon permit holders are more likely to hold only one permit
(Fig. \ref{fig:ts}C). Despite making up only 10--30% of the individuals,
multiple permit holders make approximately half of the overall revenue (Fig.
\ref{fig:ts}B). The mean effective fished species diversity of this revenue
declined somewhat from 1980--2000 but has since remained steady (Fig.
\ref{fig:ts}D). However, this trend appears to be driven by an increase in
salmon specialists in the 1980s and 1990s --- the trend disappears if salmon
license holders are excluded (Fig. \ref{fig:ts}D).

@@

Comparing single permit holders to individuals holding that same permit plus
additional permits illustrates the expected reduction in revenue variability
from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
S\ref{fig:spoke-all}). For example, individuals fishing a halibut longline
permit plus one additional permit always had lower expected revenue variability
compared to individuals fishing only a halibut longline permit (median reduction
of `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
\ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet strategy
substantially reduced expected revenue variability (Fig. \ref{fig:spokes}B).
Adding permits also usually increased expected revenue, although the permits
themselves represent a significant expense (e.g. TODO).

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given permit set is more variable (Fig.
\ref{fig:within}). After accounting for fishing effort (Fig. S\ref{fig:fe}),
specializing (decreasing species diversity) tends to be associated with
increased revenue (~0.5% increased revenue per 1% decrease in species diversity
on average or TODO ... `r g2_dens`% probability; Fig. \ref{fig:within}). This is
particularly clear for permits where individuals tend to specialize on sockeye
salmon. For example, individuals with setnet permits for salmon in Prince
William Sound, Cook Inlet, Kodiak, Bristol Bay, or the Alaska Peninsula target
primarily sockeye salmon and experience a ~1.5% increase in revenue for a 1%
decrease in fished species diversity (Fig. \ref{fig:within}B). The converse is
also usually true --- generalizing tends to be associated with lower expected
revenue after accounting for fishing effort (i.e. Fig. \ref{fig:within}A and
\ref{fig:within}B are near mirror images left to right). This benefit of
specializing is only apparent after accounting for fishing effort. Given that
fishing for more species tends to be associated with increased fishing effort
(fishing more days), generalizing tends to be associated with higher revenue
when not standardizing for days finished (Fig. S-TODO).

After accounting for fishing effort, adjusting (increasing or decreasing) an
individual's fished species diversity from year-to-year within the constraints
of a permit set tends to be associated with greater expected variability or
uncertainty in year-to-year revenue (Fig. \ref{fig:within} strategies above 0;
Fig. S\ref{fig:fe}). This effect was strongest for individuals with a halibut
longline permit or halibut and sablefish longline permits. Individuals with
these strategies experienced ~2% increase in standardized variability for a 1%
change in fished species diversity (Fig. \ref{fig:within}). Levels of fished
species diversity levels above one for these permits are primarily due to
bycatch.

The Alaskan halibut and sablefish fisheries were converted to an individual
fishery quota (IFQ) system in 1994. This catch share system... TODO. We refit
our models allowing all permit-level effects to vary separately before and after
the introduction of IFQs for strategies involving halibut and sablefish permits.
The introduction of IFQs reduced variability by ~25% (XX--XX 95% CI) for
individuals with only halibut or sablefish permits (Fig. S\ref{fig:ifq}A).
Furthermore, for most strategies, the effect of increased revenue variability
associated with changing species diversity became stronger after the
introduction of IFQs (Fig. S\ref{fig:ifq}B). This was especially true for
individuals fishing only halibut or sablefish and less so for individuals
combining halibut or sable fish with a salmon permit (Fig. S\ref{fig:ifq}B).

@@

\caption{ Trends in revenue and diversification over time. (A) Map of fishing
regions in Alaska, USA. (B) Total revenue in billions of 2009 U.S. dollars from
`r range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue)
permit holders. (C) Proportion of single permit holders for all fisheries (thick
black line) and with permit holders omitted each year who fished any of the `r 
n_jacknifed` top species groups (thin lines). (D) Mean effective diversity of
species fished (Simpson's diversity index) for all fisheries (thick black line)
and with permit holders omitted each year who fished any of the `r n_jacknifed`
top species groups (thin lines). } \label{fig:ts} \end{center} \end{figure}

@@

\caption{Diversifying permits reduces risk and usually increases revenue. Two
examples of expected revenue variability and median revenue for owners of (A)
individual halibut and (B) gillnet herring roe permits (dark blue) compared with
owners of the same permit plus one more (medium blue) or two more permits (light
blue). On the y-axis, dots and lines represent posterior median and 95\%
credible intervals of expected standard deviation.}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1.1\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}

\caption{Effects of (A) generalizing and (B) specializing from year-to-year for
individual permit holders \emph{within} the `r nstr` most common strategies.
Changing species diversity from year-to-year is associated with higher
variability for individuals within the strategies above the zero line.
Generalizing (A) or specializing (B) is associated with greater expected revenue
to the right of the zero line and lower expected revenue to the left of the zero
line. These effects are estimated holding fishing season length constant from
year-to-year. Dots and line segments represent medians and 50\% credible
intervals of posteriors. }

@@

>>>>

Managing natural resources and sustaining the livelihoods of those that depend
on their extraction is hard. These professions tend to have high year-to-year
variability in income relative to other professions [@mishra2002; @sethi2012b].
In addition to economic risk, such as changes in demand or prices, individuals
dependent on natural resources are also subject to biological and environmental
variability. For example, drought and flooding are a major source of risk for
agricultural food security and farmers' incomes [@gregory2005], and catastrophic
disease outbreaks and wildfires increase risk for the logging industry
[@kirilenko2007].

Individuals who rely on natural resources for income develop strategies to
reduce income variability. For example, farmers may rotate their crops to buffer
short-term climate variation and long-term environmental trends [@zentner1988].
Similarly, loggers maintain a diverse mix of tree species, age structure, and
harvest in geographical areas to buffer against disease outbreak, wildfires, and
price variation (REF). However, external forces may restrict the available
options to dampen income variability. Otherwise well-intentioned regulations may
limit how individuals diversify, or incentivize against diversification. For
example, in terrestrial agriculture, crop subsidies in the US may incentivize
some farms to specialize (e.g. dairy, grains, soybeans; @mishra2002). In other
instances, external market forces (e.g. prices or demand) may discourage the
adoption of diverse strategies and help promote specialization [@bradshaw2004].

Commercial fishers present an interesting example of diversification, given the
extreme risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
obtainable by commercial fishers [@kasperski2013], possibly because both fish
and fishers are spatially mobile. However, fishers still have the ability to
lower risk. They can, for example, target as many species as permitted by
a license, own multiple licenses, or diversify spatially and fish in multiple
regions.

The effectiveness of these diversification strategies for increasing revenue and
reducing revenue variability remains uncertain. For example, theory and
empirical studies suggest there can be substantial benefits to specialization
for both farmers and fishers [@abdulai2000; @arnason2005]. Furthermore, the
majority of research examining the effects of diversification on revenue do so
at aggregate levels that combine individuals (e.g. households, counties, or
communities; [@kasperski2013; @sethi2014]). Because of this, key uncertainties
remain about the fundamental relationship between diversity, fishing revenue,
and fishing revenue variability at the level of the individual.

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability. Fisheries in Alaska are among the highest-valued fisheries in the
world with ~2.7 billion kg landed biomass valued at nearly 2 billion USD
annually [@nmfs2016]. Like much of the rest of the world, the days of open,
unregulated fisheries in most developed countries have passed --- fishers must
own a license to participate in virtually all fisheries [@hilborn2003a]. And
while such regulations have helped reduce overfishing, they limit the available
strategies to reduce income variability. For example, many permits limit fishing
to a single species, diversifying by purchasing additional fishing permits can
be expensive (e.g. ~\$192K USD for a king and tanner crab permit in Southeast
Alaska in 2016; <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>), and some
fisheries limit the entry of new participants through individual fishing quotas
(IFQs) [@grimm2012].

Using a unique dataset collected over 30 years (1985--2014) on individuals,
rather than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals rather than
vessels we can track people that switch fishery permits over time or use
multiple vessels. Similarly, we are able to disentangle fishers that share the
same vessel. We develop and apply a hierarchical Bayesian model to understand
the effect of diversity on both individual revenue and revenue variability that
builds on results from vessel [@kasperski2013], fleet [@sethi2012b], and
community-level [@sethi2014] variability. Our model controls for fishing effort
and latent time effects (e.g. changing environmental conditions), simultaneously
estimates the effect of species diversity on individual revenue and revenue
variability, and estimates effects both within and across fishing strategies
(sets of similar licenses or permits). To our knowledge, this is the first
application of this type of model in any natural resource field.

Our hierarchical variance-function regression model [@davidian1987; @hall1989]
simultaneously models changes in the mean and variance of individual revenue as
a function of covariates. Whereas traditional hierarchical linear models assume
homoscedasticity --- that the residual variance does not vary systematically ---
the variance-function regression model allows the variance to be heteroscedastic
and vary as a function of predictors. Our specific model allows the revenue rate
of return (log of the ratio of year-to-year revenue) and the revenue variability
(the standard deviation of this revenue rate of return) to vary by fishing
strategy (fishing permit sets), fishing effort (season length), and levels of
species diversification.

@@

We first evaluated trends in revenue and diversification over time for
individuals fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A).
The majority (70--90%) of fishers hold single permits, with a slight increase in
single-permit holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is
most sensitive to halibut or salmon permit holders --- compared to other
fisheries, halibut permit holders are more likely to hold halibut plus other
permits, whereas salmon permit holders are more likely to hold only one permit
(Fig. \ref{fig:ts}C). Despite making up only 10--30% of the individuals,
multiple permit holders make approximately half of the overall revenue (Fig.
\ref{fig:ts}B). The mean effective fished species diversity of this revenue
declined somewhat from 1980--2000 but has since remained steady (Fig.
\ref{fig:ts}D). However, this trend appears to be driven by an increase in
salmon specialists in the 1980s and 1990s --- the trend disappears if salmon
license holders are excluded (Fig. \ref{fig:ts}D).

@@

Comparing single permit holders to individuals holding that same permit plus
additional permits illustrates the expected reduction in revenue variability
from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
S\ref{fig:spoke-all}). For example, individuals fishing a halibut longline
permit plus one additional permit always had lower expected revenue variability
compared to individuals fishing only a halibut longline permit (median reduction
of `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
\ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet strategy
substantially reduced expected revenue variability (Fig. \ref{fig:spokes}B).
Adding permits also usually increased expected revenue, although the permits
themselves represent a significant expense (e.g. ~\$192K USD for a king and
tanner crab permit in Southeast Alaska in 2016;
<https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>).

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given permit set is more variable (Fig.
\ref{fig:within}). After accounting for fishing effort, specializing tends to be
associated with increased revenue (Fig. \ref{fig:within}; Fig. S\ref{fig:fe}).
On average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
decrease in species diversity, and across strategies, `r g2_dens`% of the
probability density supports a positive relationship between specialization and
revenue. The converse is also usually true --- generalizing tends to be
associated with lower expected revenue after accounting for fishing effort (i.e.
Fig. S\ref{fig:mirror}A and \ref{fig:mirror}B are near mirror images left to
right). This benefit of specializing is only apparent after accounting for
fishing effort. Given that fishing for more species tends to be associated with
increased fishing effort (fishing more days), generalizing tends to be
associated with higher revenue when not standardizing for days finished (Fig.
S-TODO).

After accounting for fishing effort, adjusting (increasing or decreasing) an
individual's fished species diversity within the constraints of a permit set
tends to be associated with greater expected variability or uncertainty in
year-to-year revenue (Fig. \ref{fig:within} strategies above 0; Fig.
S\ref{fig:fe}). This effect was strongest for individuals with a halibut permit
or halibut and sablefish permits. Individuals with these strategies experienced
~2% increase in variability for a 1% change in the year-to-year ratio of species
diversity (Fig. \ref{fig:within}). Levels of fished species diversity above one
for these particular permits are primarily due to bycatch.

The management of the Alaskan halibut and sablefish fisheries was converted to
individual fishery quotas (IFQ) in 1994. This catch share system... TODO. We
refit our models allowing all permit-level effects to vary separately before and
after the introduction of IFQs for strategies involving halibut and sablefish
permits. The introduction of IFQs reduced variability by ~25% (XX--XX 95% CI)
for individuals with only halibut or sablefish permits (Fig. S\ref{fig:ifq}A).
Furthermore, for most strategies, the effect of increased revenue variability
associated with changing species diversity became stronger after the
introduction of IFQs (Fig. S\ref{fig:ifq}B). This was especially true for
individuals fishing only halibut or sablefish and less so for individuals
combining halibut or sable fish with a salmon permit (Fig. S\ref{fig:ifq}B).

@@

\caption{Trends in revenue and diversification over time. (A) Map of fishing
regions in Alaska, USA. (B) Total revenue in billions of 2009 U.S. dollars from
`r range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue)
permit holders. (C) Proportion of single permit holders for all fisheries (thick
black line) and with permit holders omitted each year who fished any of the `r 
n_jacknifed` top species groups (thin lines). (D) Mean effective diversity of
species fished (Simpson's diversity index) for all fisheries (thick black line)
and with permit holders omitted each year who fished any of the `r n_jacknifed`
top species groups (thin lines). } \label{fig:ts} \end{center} \end{figure}

@@

\caption{Diversifying permits reduces risk and usually increases revenue. Two
examples of expected revenue variability and median revenue for owners of (A)
individual halibut and (B) gillnet herring roe permits (dark blue) compared with
owners of the same permit plus one more (medium blue) or two more (light
blue) permits. On the y-axis, dots and lines represent posterior median and 95\%
credible intervals of expected standard deviations of year-to-year revenue rates
of return.}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\textwidth]{../figs/stan-offset-break-anti-spaghetti-specializing.pdf}

\caption{Effect of specializing on year-to-year changes in revenue and revenue
variability for individual permit holders \emph{within} the `r nstr` most common
strategies. Changing species diversity from year-to-year is associated with
higher variability for individuals within the strategies above the zero line.
Specializing is associated with greater expected revenue to the right of the
zero line and lower expected revenue to the left of the zero line. For example,
individuals with setnet permits for salmon in Prince William Sound, Cook Inlet,
Kodiak, Bristol Bay, or the Alaska Peninsula target primarily sockeye salmon and
experience a $\sim$1.5\% increase in their year-to-year revenue ratio for a 1\%
decrease in their year-to-year species diversity. These effects are estimated
holding fishing season length constant from year-to-year. Dots and line segments
represent medians and 50\% credible intervals of posteriors. The effect of
generalizing (increasing species diversity) is similar (Fig. S\ref{fig:mirror})}.

@@

# Materials and methods

We used combinations of state and federal fishing permits held by an individual
in in a specific year to define fishing "strategies", representing the species
targeted (Supporting Information). For example, we combined permits that were
otherwise the same except for vessel size (e.g. a longliner fishing sablefish
on a vessel under 60 ft was considered to have the same strategy as a longliner
fishing sablefish on a vessel over 60 ft), and we combined a number of permits
that were targeting a single species and differed only in region fished.
Therefore, for example, someone fishing herring roe in Kodiak and someone
fishing herring roe in Alaska Peninsula were both considered to have a "herring
roe" permit group. For salmon permits, we formed salmon strategies based on
similarities in gear type and species compositon (Supporting Information).

\begin{equation}
\log(R_{i,t}) \sim \mathrm{N} \left(\mu_{i,t}, \sigma_{i,t}^2 \right).
\end{equation}

\noindent Each year $t$ of revenue for fisher $i$ is assigned a strategy (permit
set) indexed by $j$ (Supporting Information, Table \ref{tab:strategies}) and
$\mu_{i,t}$ is modelled as:

\begin{equation}
\label{eq:mean}
\mu_{i,t} = \beta_{0,j,t} -
  \beta_{1,j} S_{i,t} I_{i,t} +
  \beta_{2,j} S_{i,t} I_{i,t} +
  \beta_3 D_{i,t} -
  \beta_4 S_{i,t} D_{i,t} I_{i,t} +
  \beta_5 S_{i,t} D_{i,t} I_{i,t} +
  \log(R_{i,t-1}),
\end{equation}

\noindent where $S_{i,t}$ is the log ratio of species diversity from
year-to-year, $\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$,
$D_{i,t}$ is the log ratio of days fished from year-to-year, $\log
\left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and $I_{i,t}$ is an
indicator variable that takes a value $0$ if $\mathrm{div}_{i,t} \ge
\mathrm{div}_{i,t-1}$ (species diversity increasing; generalizing) and takes
a value $1$ if $\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$ (species diversity
decreasing; specializing). Estimating a separate coefficient for the effect of
increasing diversity and decreasing diversity allows the costs and benefits of
these changes to be asymmetric.

We use $\log(R_{i,t-1}$) as an offset, and thus our model describes the log
ratio of revenue from year-to-year, i.e., $\log(R_{i,t}) - \log(R_{i,t-1})
= \log(R_{i,t}/R_{i,t-1})$. We model change in revenue because preliminary
analysis revealed non-stationary trends in revenue (resulting from experience or
other unmodeled factors). This approach eliminates having to deal with the
non-stationary time series and eliminates having to estimate thousands of
mean-revenue intercepts associated with individual fishers. We removed
person-year combinations without a subsequent year of revenue.

The $\beta$ terms represent estimated coefficients, some of which are allowed to
vary as "random effect" terms. The intercepts $\beta_{0,j,t}$ are allowed to
vary by strategy-year combinations, are centred on zero, and are constrained by
a normal distribution:

\begin{align}
\beta_{1,j} &\sim \mathrm{N} \left(\mu_{\beta_1}, \tau^2_{\beta_1}\right),\\
\beta_{2,j} &\sim \mathrm{N} \left(\mu_{\beta_2}, \tau^2_{\beta_2}\right).
\end{align}

\noindent The $\beta_{1,j}$'s represent how much a 1\% decrease in species
diversity within a strategy translates to a Y\% change in revenue. The
$\beta_{2,j}$'s represent how much a 1\% increase in species diversity within
a strategy translates to a Y\% change in revenue. These are after controlling
for changes in effort, $\beta_3$, and the interaction between changes in effort
and changes in species diversity, $\beta_4$ and $\beta_5$.

\begin{equation}
\label{eq:cv}
\sigma_{i,t} =
\exp \left( \gamma_{0,j} -
\gamma_{1,j} S_{i,t} I_{i,t} +
\gamma_{2,j} S_{i,t} I_{i,t} +
\gamma_3 D_{i,t} -
\gamma_4 S_{i,t} D_{i,t} I_{i,t} +
\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
\end{equation}

\begin{align}
\gamma_{0,j}  &\sim \mathrm{N} \left(\eta_0 + \eta_1 M_j + \eta_2 E_j, \tau^2_{\gamma_0}\right),\\
\gamma_{1,j}  &\sim \mathrm{N} \left(\mu_{\gamma_1}, \tau^2_{\gamma_1}\right),\\
\gamma_{2,j}  &\sim \mathrm{N} \left(\mu_{\gamma_2}, \tau^2_{\gamma_2}\right).
\end{align}

\noindent The mean of the $\gamma_{0,j}$ coefficients is modelled with two
strategy-level predictors. The symbols $M_j$ and $E_j$ represent the mean
species diversity and mean combined fishing season length (effort) for strategy
$j$. The coefficients $\eta_1$ and $\eta_2$ represent predictors in this
strategy-level regression estimating effects across strategies.

We fit our models with Stan 2.11.1 [@hoffman2014; @standevelopmentteam2015;
@standevelopmentteam2015a; @carpenter2015] and R 3.3.1 [@rcoreteam2016]. Stan
implements the No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm
[@hoffman2014] to perform Bayesian statistical inference. We assigned weakly
informative priors on all parameters: $N(0, 2^2)$ priors on all $\beta$,
$\gamma$, and $\eta$ parameters, and half-$t(3, 0, 2)$ priors (i.e. degrees of
freedom of 3 and scale of 2) on all $\sigma$ and $\tau$. We ran four chains and
2,000 iterations and discarded the first 1,000 iterations of each chain as warm
up. We checked for chain convergence visually with trace plots, ensured that
$\hat{R}<1.05$ (the potential scale reduction factor) and that the effective
sample size, as calculated by the rstan R package [@standevelopmentteam2015],
was greater than 200 for all parameters [@gelman2014]. We verified that our
model returned sensible estimates by comparing our estimates to an ad-hoc
two-stage model fit with the lme4 R package [@bates2015], where we first fit
a mixed-effect regression model to the mean, and then modelled the absolute
residuals from that first model in a second mixed effects model representing the
variance component. We also validated the model with simulated data, and were
able to recover the true values (Supporting Information). Code used in our
analysis is available at <https://github.com/NCEAS/pfx-commercial/portfolios/>.

*TODO: move the following into the results section or into the methods section:*
